---
title: "Homework 2"
subtitle: "DSCI 101"
author: "Your Name Here"
date: "Due 02-02-2024"
output: html_document
---
Homework should be written as an R Markdown file and the knitted output should be submitted. Homework should be turned in on Sakai.

1. (4pts) Clone this repo into RStudio (if you haven't done so yet): https://github.com/wabdalla/DSCI_101


Take a screenshot of your RStudio. You will get full credit if in the screenshot I see the class folders (under the files tab) and
I see the git tab on the upper right window. For further instructions on how to include pictures, see the R Markdown cheat sheet. 

2. (3pts) Create a numeric vector containing the values 5, 10, 15, 20, and 25. Calculate the mean of this vector and assign it to a variable called `mean_value`. Print the `mean_value`

```{r ex2}


```


3. (3pts) Create a data frame with the following information about three of your friends:

* Name (character)
* Age (numeric)
* Favorite Color (character)

Make sure to include at least three rows of data. Print the entire data frame. Name it whatever you like.

```{r ex3}

```

4. (3pts) Create two numeric vectors: `vector1` with values 1, 2, 3, and `vector2` with values 4, 5, 6. Perform the following operations:

* Add `vector1` and `vector2` and store the result in a new vector called `sum_vector.`
* Subtract `vector2` from `vector1` and store the result in a new vector called `diff_vector.`
* Multiply `vector1` and `vector2` element-wise and store the result in a new vector called `prod_vector.`
* Divide `vector1` by `vector2` element-wise and store the result in a new vector called `quot_vector.`

Print all four result vectors.

```{r ex4}

```

5. (12pts) Using the `top_colleges_2022` dataset answer the following questions.

a) Read in the csv file and name the variable `college`. This should be simple if question 1 has been completed as the github will download the data folder.

```{r ex5a}

```

b) How many colleges in this data set are in Illinois.

```{r ex5b}

```

c) Find the undergraduate student population of LUC. How man colleges (in this dataset) have a larger undergraduate student population than LUC?

```{r ex5c}

```

