---
title: "Homework 5"
author: "Your Name Here"
date: "2024-03-12"
output: html_document
editor_options: 
  chunk_output_type: console
---


### **Instructions**

Use functions learned from the tidyverse package to complete this homework assignment. Load all the packages you will use throughout this homework in the loading packages section below.


Submit the html file in Sakai before the deadline.


### **Loading Packages**

(5 points)

```{r packages, message=FALSE}


```


### **Exercise 1**

Using the `Teams` data frame in the Lahman package:

a) (10 points) Make a line plot of total home runs per season and stratify (facet) by league. Do you see anything interesting that you observe (per league) throughout the years? Note: a season is an entire year. 

*YOUR WRITTEN ANSWER HERE*

```{r ex1a, message = FALSE}




```


b) (10 points) Create an indicator variable called "winning record" which is defined as TRUE if the number of wins is greater than the number of losses and FALSE otherwise. Plot a scatter plot of Runs (R) (in x axis) vs Runs against (RA) (in y axis) with the color of each point showing whether that team had a winning record or not.


```{r ex1b, message = FALSE}


```

c) (5 points) Create a scatterplot of strikeouts (SO) (x axis) vs walks (BB) (y axis) faceting by Ranks 1 through 5. 


```{r ex1c, message = FALSE}


```

d) (10 points) Create a categorical variable with 4 levels: before 1900, 1900-1949, 1950-1999, 2000 or later. (Hint: There is a function called 'cut' discussed in class that will help you here. Make sure your breakpoints correctly include the years in each category). Use this as the color of the point in the scatterplot of strikeouts (x axis) vs walks (y axis) that you created in the previous problem. Now what do you see?

*YOUR WRITTEN ANSWER HERE*

```{r ex1d, message = FALSE}



```


### **Exercise 2**

Using the `flights` data frame in the nycflight13 package:

a) (5 points) Create a histogram of the departure delays times (dep_delay) for all flights that had a delay time greater than 0.


```{r ex2a, message = FALSE}


```

b) (5 points) Repeat the same histogram, but now facet by month. 


```{r ex2b, message = FALSE}



```

c) (10 points) Create side-by-side boxplots of delay times for flights. Create categories - if a flight was scheduled to leave before noon or afternoon (use 1200 for the time). Only select the top 5 airlines (in terms of number of flights).


```{r ex2c, message = FALSE}


```

### **Exercise 3**

Using `top_colleges_2022` recreate these images.

a) (10 points) Recreate Plot 1 from the attached pdf file. Use `firebrick1` and `dodgerblue1` colors to recreate the plot.


```{r ex3a, message = FALSE}



```




b) (20 points) Recreate Plots 2 & 3 and report what stands out from each one.

*YOUR WRITTEN ANSWER HERE*

```{r ex3b, message = FALSE}




```


### **Exercise 4**

(15 points) Suppose you have 5 students in a classroom. Let the vector `stud_name` represent these students' names. Let vectors `exam_1` and `exam_2` be two vectors with their respective scores in Exams 1 & 2. 

```{r ex4info, message = FALSE}
stud_name <- c("Emily", "John", "Steve", "Mike", "Sarah")

exam_1<-c(90, 50, 45, 52, 100)

exam_2<-c(100, 50, 40, 40, 100)
```
A student passes if they have an average score of 50. They fail if they have their average score is below 50. And they will be granted an extension if their average score is exactly 50. 

Create a for loop combined with if-else statements that computes each student's average score in the class. 
Within the for loop, it should be printed whether each student passed, failed, or will be granted an extension.
The final output should look like: 
"Emily passed the class with an average of 95."
"John will be granted an extension in the class with an average of 50"
"Steve failed the class with an average of 42.5"... etc.

```{r ex4, message = FALSE}


```