---
title: "Homework 4"
author: "YOUR NAME HERE"
date: "2024-03-13"
output: html_document
editor_options: 
  chunk_output_type: console
---

### **Instructions**

Use functions learned from the tidyverse package to complete this homework assignment. Load all the packages you will use throughout this homework in the loading packages section below.


Submit the html file in Sakai before the deadline.


### **Loading Packages**

(5 points)

```{r packages, message=FALSE, warning = FALSE}

```

### **Exercise 1**

Use the `Batting`, `Pitching`, and `People` dataframes in the Lahman package to answer the following questions. Make sure that you read the background information on these dataframes and understand them well before answering each question. 

a) (20 points) Name every player (first and last name) in baseball history who has accumulated across their careers 300 home runs (HR) or more AND 300 stolen bases (SB) or more. You can find the first and last name of the player in the `People` data frame. Join this to your result along with the total home runs and total bases stolen for each of these elite players.
***Your final output should be a printed dataframe that has four columns reflecting the players' first name and last name and the total homeruns and total stolen bases***


```{r ex1a, message = FALSE}

```


b) (20 points) Similarly, name every pitcher (first and last name) in baseball history who has accumulated 300 wins (W) or more AND 4,000 strikeouts (SO) or more. 
***Your final output should be a printed dataframe that has four columns reflecting the players' first name and last name and the total wins and total strikeouts***

```{r ex1b, message = FALSE}

```

c) (20 points) Identify the name of the players who have hit 50 home runs or more in a single season. Of those players, which player had the lowest batting average (hits/at bats [H/AB])? ***Your final output should be a printed dataframe that has five columns and one row reflecting the one player's first name, last name, the year of the season, the number of homeruns, and the batting average (H/AB)***

```{r ex1c, message = FALSE}

```

### **Exercise 2**

Use the `nycflights13` package and the `flights` and `planes` tables to answer the following questions. Make sure that you read the background information on these dataframes and understand them well before answering each question. 

a) (15 points) What is the oldest plane (specified by the tailnum variable) that flew from New York City airports in 2013 and how many flights did this plane conduct from New York City airports in 2013? ***Your final output should be a printed dataframe that has three columns reflecting the tailnumber of the plane, the year it was manufactured, and the number of flights this plane conducted from NYC***


```{r ex2a, message = FALSE}

```

b) (10 points) How many ***different*** airplanes (specified by the tailnum) are in the `flights` dataset? How many ***different*** airplanes (specified by the tailnum) are in the `planes` dataset? 

*YOUR WRITTEN ANSWER HERE*



```{r ex2b, message = FALSE}

```

c) (10 points) In 2013, there were 336,776 flights that left New York City, how many of those airplanes are included in the planes table? What does this imply if we do an inner_join with the `flights` and `planes` dataset by `tailnum`? (You don't need to do an inner join, simply explain the implication).

*YOUR WRITTEN ANSWER HERE*


```{r ex2c, message = FALSE}

```