---
title: "Homework 3"
author: "Your Name Here"
date: "2024-02-16"
output: html_document
editor_options: 
  chunk_output_type: console
---



### **Instructions**

Use functions learned from the tidyverse package to complete this homework assignment. Load all the packages you will use throughout this homework in the loading packages section below. 


For Exercise 2, you will use the `Lahman` package. Please install the package `install.packages("Lahman")` in your console, then load the package in the loading packages section below. 

Submit the html file in Sakai before the deadline.


### **Loading Packages**

(5 points)

```{r packages, message=FALSE}

```


### **Exercise 1**

Using the `cubs_all_time` dataset found in the data folder. Answer the following questions using R:

a) (10 points) How many total games have the Cubs won between 1876 and 2022 and how many total games have the
Cubs lost between 1876 and 2022? ***Your final output should be a printed dataframe that has two columns reflecting these two numbers.***

*YOUR WRITTEN ANSWER HERE*


```{r ex1a, message = FALSE}


```

b) (10 points) What year did the Cubs score the most runs? What year did the Cubs score the fewest runs? Do you have any thoughts about the year that the Cubs scored the fewest runs? ***Your final output should be a printed dataframe that has two columns representing the year and the number of runs.***

*YOUR WRITTEN ANSWER HERE*


```{r ex1b}


```

c) (10 points) In how many seasons was the Cubs total attendance (i.e. the variable Attendance) over 3 million? ***Your final output should be a printed dataframe that has a column representing the number of seasons.***

*YOUR WRITTEN ANSWER HERE*



```{r ex1c}



```

d) (10 points) In the years between 1982 and 2000 (inclusive), what was the greatest ratio of runs for (R) to runs against (RA) in a single season? ***Your final output should be a printed dataframe that has a column representing the greatest ratio.***

*YOUR WRITTEN ANSWER HERE*


```{r ex1d}


```

### **Exercise 2**

The data you will use in this exercise is called `Teams` and you will find it inside the Lahman package. If you loaded the package at the beginning of this homework, the dataset will automatically be loaded. To put this dataset in your environment simply type `data("Teams")` in your codes. Use `?Teams` or `help(Teams)` to get a description of this data. 


a) (10 points) Create a data frame that is a subset of the Teams data frame that contains only the years from 2000 through 2009 (inclusive) and the variables yearID, W, and L. ***Do not print your dataframe.***

```{r ex2a}



```


b) (10 points) How many years did the Chicago Cubs (teamID is "CHN") hit at least 200 HRs in a season and what was the median number of wins in those seasons (the seasons with at least 200 HRs). ***Your final output should be a printed dataframe that has two columns representing the number of seasons and the median number of wins.***


*YOUR WRITTEN ANSWER HERE*



```{r ex2b}

```

### **Exercise 3**

The data you will use in this exercise is called `flights` and you will find it inside the `nycflights13` package. If you loaded the package at the beginning of this homework, the dataset will automatically be loaded. To put this dataset in your environment simply type `data("flights")` in your codes. Use `?flights` or `help(flights)` to get a description of this data.


a) (15 points) What month of the year had the highest average time of departure delayed flights and what is the value of this average? Note: Make sure you do not account for missing values when computing the averages. ***Your final output should be a printed dataframe that has two columns representing the month and the average time of departure. The final dataframe must only contain one row showing the month with the respective average delay.***


*YOUR WRITTEN ANSWER HERE*



```{r ex3a}


```

b) (20 points) Which origin airport has the highest proportion of delayed flights leaving? Which origin airport has the highest proportion of flight arriving late? Note: a flight is considered delayed if the delayed time is greater than 0 minutes, otherwise (including missing values) it is not delayed. ***Your final output should be a printed dataframe that has three columns representing the origin airport, the proportion of delayed departure flights, and the proportion of delayed arrival flights.***


*YOUR WRITTEN ANSWER HERE*


```{r ex3b}


```

