---
title: "Homework 8"
author: "YOUR NAME HERE"
date: "2024-12-04"
output: html_document
editor_options: 
  chunk_output_type: console
---


### **File in html Format**

(5 points)

Submit the html file in Sakai before the deadline.



### **Loading Packages**

(5 points)

```{r packages, message=FALSE, warning = FALSE}



```


### **Exercise: Sampling Distribution & Bootstrapping**


In 2003 a study was done to see how long medical residents worked per week. Suppose the number of hours worked per week by medical residents is normally distributed with a mean of 81.7 and standard deviation of 6.9 hours. 


a) (10 points) Based on the scenario above, create a hypothetical population. Assume a population of N = 5000000. ***Do not print your results.***

```{r, ex1a}


```
b) (10 points) Collect ***one*** sample from the population you created in part (a). Sample size = 500. Save this sample as a vector called one_sample. ***Do not print your results.***

```{r, ex1b}



```

c) (20 points) Create a *Sampling Distribution of the* ***median***, with a sample size of 500. To create this sampling distribution, first create a function that randomly selects a sample of size n from a population vector and computes the median of the sample. Arguments should be: i, a population vector, and sample size. The return should be the median of the sample. Second, once you build the function, use a the map function to randomly select multiple samples of size n = 500 from the population vector in part (a). Iterate 10000 times. Meaning randomly draw 10000 samples, each one of size n = 500. ***Do not print your results.***

```{r, ex1c}



```

d) (10 points) Use your results from part c to plot your sampling distribution of the median (with a sample size of n = 500). ***Make sure your plot is printed when knitting the file*** 

```{r, ex1d, message = FALSE}



```


e) (10 points) Compute the standard deviation of the sampling distribution you created in part (c). This is the standard error of your sampling distribution. ***Print your results***.   

```{r, ex1e}



```


f) (20 points) Use the bootstrap to estimate the standard error of the sampling distribution you created in the previous steps. Use the sample you created in part (b). Use 10,000 iterations. ***Print your estimated standard error.***  

```{r, ex1f}




```

g) (10 points) Compare your results you got in part e) vs part f) (subtract the values). What are your thoughts on this value?

*YOUR WRITTEN ANSWER HERE*

```{r, ex1g}



```









