---
title: "Second Midterm In Class"
author: "YOUR NAME HERE"
date: "2026-03-30"
output: html_document
---

Same rules from the take home exam hold.

### **Loading Packages**

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
```

### **Read in Datasets**

Read in `midterm_data.csv` and name it `midterm_data`.

```{r read_data, message = FALSE, warning = FALSE}



```


### **Question 1** 

How many students are there per course section and graduating class? 

The output in one column should look like: 

DSCI 101 - 002 - Freshman 

DSCI 101 - 002 - Junior     

DSCI 101 - 002 - Senior

... etc. 

***Your final output should be a printed dataframe with 2 columns and 11 rows. ***

```{r question_1, message = FALSE, warning = FALSE}




```



### **Question 2**

a. Transform the `birthdate` column into a variable of class "Date" (make this transformation within the same dataframe and make sure the column is still named `birthdate`).

```{r question_2a, message = FALSE, warning = FALSE}




```
b. In what month were most students born? To get full credit, make sure the full name of the month (i.e., "January") appears. 


***Your final output should have 2 columns and one row.***

```{r question_2b, message = FALSE, warning = FALSE}





```


### **Question 3**

Recreate Question 3 Dataframe.

Question 3 Dataframe represents the number of students per graduating class vs their belief on who would win a fight (Grilled Cheese, Hot Dog, Taco). Use `midterm_data` to repoduce this dataframe.

Hint: You need to compute the number of students by per fight and graduating class.   

***Your final output should be the recreated dataframe printed.*** 

```{r question_3, message = FALSE, warning = FALSE}




```


### **Question 4**

Recreate Question 4 Plot. 

***Make sure the plot is printed***

```{r question_4, message = FALSE, warning = FALSE}





```


### **Question 5**

The function below takes the square of a number that is larger than 205 and takes the cube otherwise. 
Apply this function to each observation of the `minutes_screen` column using one of the `map` functions. After that, take the average across all these observations. 

***You final output should be one printed value.***

```{r question_5, message = FALSE, warning = FALSE}


square_cube <- function(x){
  if(x>205){
    output = x^2
  }else{
    output = x^3
  }
}






```
