---
title: "Exam 2"
author: "YOUR NAME HERE"
date: "2025-03-30"
output: html_document
editor_options: 
  chunk_output_type: console
---


### **Instructions**

Below is your second midterm take home exam for DSCI 101 - Spring 2025. This exam should be written only by the student. By taking this exam you are not working with any other student in the class or other resources outside of this class. 

You are allowed to consult: 

* Class notes.
* Class R Scripts. 
* Class homework. 
* Help buttons in R. 

You are NOT allowed to: 

* Search for answers using search engines such as Google. 
* Search for answers using artificial intelligence programs such as ChatGPT.
* Consult with anyone (i.e., classmates, tutors, friends, family members... etc) 

**If you are found to be working with someone else (classmate, tutor... etc.) or have used AI softwares to produce your answers, an automatic 0 will be given on the midterm exam (both in class and take home) and you will be reported to the Dean's Office.**

Use functions learned from the `tidyverse` package to complete this exam. Load all the packages you will use throughout this exam in the loading packages section below.

Answer each question using this same file. Your codes should go inside the codechunks. 

For Question 5, you will need to recreate Plot # 1, which is a pdf that is included in the same folder. 

### **Dataset Description: Student Data**

You will be examining the class survey dataset (`midterm_data`) that you had to take at the beginning of the semester. I have cleaned the dataset for general use and added a few columns with "fake" data. Please refer back to the survey [here](https://forms.office.com/Pages/ResponsePage.aspx?id=408fApwrJEiDeLvPnsWsy0GW1-51jWZMpmlDitVq_wZUNDdUS09YSkQ1RUFRSzVCWkVET1k3NU9GNS4u). 

In addition to the class survey, you will need to examine three more datasets: `chicago_covid_cases` (see description of this data below), `midterm_studentID`, and `colors_data` that are included inside the same file. 

The `midterm_studentID` dataset is a dataset that contains fake student ID's for each student in the `midterm_data` file. 

`colors_data` is a file that shows how different colors are categorized. 


### **Dataset Description: Chicago COVID Cases**

The dataset, named `chicago_covid_cases.csv`, has been sourced from the City of Chicago Data Portal. It represents COVID-19 cases in Chicago and has been cleaned to ensure a standardized format. The dataset includes key variables such as 'date,' 'total' (total cases for that particular day), 'female' (number of female cases for that particular day), 'male' (number of male cases for that particular day), and columns related to different age groups. 




### **File submitted in HTML Format **

Submit the html file in Sakai before the deadline. 10 points will be deducted from your total score if submitted in rmd format.


### **Loading Packages**


```{r packages, message = FALSE, warning = FALSE}

library(tidyverse)

```

### **Question 1**

(14 points) 


a. Read in the dataset `chicago_covid_cases.csv`. Name the dataframe `cases`.  

```{r question_1a, message = FALSE, warning = FALSE}



```

b. Transform the `date` column into a variable of class "Date" (make this transformation within the same dataframe and make sure the column is still named `date`).

```{r question_1b, message = FALSE, warning = FALSE}



```
c. Using one of the functions from the `map` family functions, check the class of all the columns in `cases`. Make sure your output is in the form of a vector and not a list. ***Your final output should be a printed vector that shows you the class of every column in the dataframe.***

```{r question_1c, message = FALSE, warning = FALSE}



```

d. Create a new column called `month` based on the `date` variable. The month column should contain the actual names of the month in a truncated form (i.e., Jan, Feb... etc). Then compute the total number of cases for each one of the 12 months. Hint: Use the total column to get the total number of cases per month. **Your final output should be a printed dataframe with 12 rows and two columns: the month and the total number of cases for each one of the months.**

```{r question_1d, message = FALSE, warning = FALSE}




```

### **Question 2**

(14 points)

a. Read in the dataset `midterm_data.csv` and save it into a dataframe called `midterm_data`.

```{r question_2a, message = FALSE, warning = FALSE}




```

b. Read in the dataset `midterm_studentID.csv` and save it into a dataframe called `midterm_studentID`. 

```{r question_2b, message = FALSE, warning = FALSE}



```

c. Using the column `studentID` in the `midterm_studentID` dataframe, calculate how many students ID have an ID that contains in it a 5. ***Your final output should be a printed dataframe with one column and one row that shows you the total number of students that have a student ID that contains a 5 in it.***

```{r question_2c, message = FALSE, warning = FALSE}



```

d. Calculate how many students have a student ID that contains the number 5 AND their favorite subject is Math. ***Your final output should be a printed dataframe with one column and one row that shows you the total number of students that have a student ID that contains a 5 in it AND their favorite subject is Math.***

```{r question_2d, message = FALSE, warning = FALSE}



```
### **Question 3**

(14 points) 

Using the `midterm_data`, make a scatter plot of student's height (inches column only) in the y axis and Year in the x axis and split the data by the students favorite color. 

Make sure that the color of the dots for each color match the color that is given, (i.e., in the legent the word Black has to match to the color "black", the word Blue has to match to the color "blue"... etc.)

Create a nice plot that has a "minimal" theme (background), an appropriate title, appropriate axis titles, and appropriate leyend title. (All titles must start with capital letters). 


```{r question_3, message = FALSE, warning = FALSE}




```



### **Question 4**

(14 points) 

a. Create a new function called color_category such that it's argument is a color_name. If the color is Blue, Yellow, or Red, then the function will return "Primary Color". If the color is Green, Orange, or Purple, then the function will return "Secondary Color". If the color is Black, White, or Brown, then the function will return "Neutral Color". Any other color or argument, then the function will return "Other Color". 

After creating the function, test the function (individually) with the color Black, Blue, Green, and Pink. 

***Note: Make sure that the first letter of the color is a capital letter (i.e. Blue).*** 


```{r question_4a, message = FALSE, warning = FALSE}



```

b. Apply the function you created in part a to the midterm_data using the `map_chr` function (or any other function in the `map` family). To accomplish this, you must create a new column by applying the new function you created in part a to the column color. Then calculate how many students like each one of the color categories. ***Your final output should be a printed dataframe with two columns and four rows that shows you the color categories and the number of students in each category.***   

```{r question_4b, message = FALSE, warning = FALSE}



```

### **Question 5**

(14 points) 

a. Read in the dataset `colors_data.csv` and save it into a dataframe called `colors_data`. 

```{r question_5a, message = FALSE, warning = FALSE}



```

b. `colors_data` is a dataframe that shows different color categories and some colors that belong to each category. Merge `colors_data` and `midterm_data` so that we can see every student's color category. ***DO NOT PRINT THE DATAFRAME.***

```{r question_5b, message = FALSE, warning = FALSE}



```

c. Using the results from part b, recreate the attached plot (Plot # 1) EXACTLY as it appears. Colors used are black, mediumblue, darkviolet, and deeppink. 

```{r question_5c, message = FALSE, warning = FALSE}




```