---
title: "Homework 2"
subtitle: "DSCI 101"
author: "YOUR NAME HERE"
date: "Due 09-22-2025"
output: html_document
---


### **Instructions**



For each exercise, write the codes inside the coding chunks provided. To avoid running into any last minute errors, I recommend knitting the file every time you finish working on an exercise. Submit the **html file** in Sakai before the deadline.


### **Exercise 1**

(12 pts) Create a numeric vector named vector_1 using the `sequence` function discussed in class. The vector should contain the values 6, 8, 10, 12, and 14. Calculate the median of this vector and save it as an object called `median_value`. ***Your final output when knitting the file should show the printed values for `median_value`***

```{r ex1, message=FALSE}



```


### **Exercise 2**

(12 pts) Create a data frame with the following information about three of your friends:

* Name (character)
* Age (numeric)
* Favorite Color (character)

Make sure to include at least three rows of data. Make sure you provide a name to this dataframe (you can name it whatever you like.) ***Your final output when knitting the file should show the printed dataframe***

```{r ex2}



```

### **Exercise 3**


(16 pts) Create two numeric vectors: `vector1` with values 1, 2, 3, and `vector2` with values 4, 5, 6. Perform the following operations:

* Add `vector1` and `vector2` and store the result in a new vector called `sum_vector.`
* Subtract `vector2` from `vector1` and store the result in a new vector called `diff_vector.`
* Multiply `vector1` and `vector2` element-wise and store the result in a new vector called `prod_vector.`
* Divide `vector1` by `vector2` element-wise and store the result in a new vector called `quot_vector.`


***Your final output when knitting the file should show the printed four result vectors.***

```{r ex3}




```

### **Exercise 4**

(28 pts) In a company, employees get different kinds of perks based on the amount of hours they worked during the year. If the employee worked less than or equal to 2000 hours, they will not get a perk and they will get paid $10 an hour. If they worked more than 2000 hours but less than or equal to 3000 hours they will get a paid trip to Hawaii and they will get paid $20 an hour. Finally, if they worked more than 3000 hours, they will get a brand new car and they will get paid $30 an hour. 
```{r ex4info, message = FALSE}
employee <- c("Mary", "John", "Tim", "Holly", "Glen")
hours <- c(2080, 3120, 2600, 1900, 4000)
```

Create a for loop that computes and prints each employee's perk along with their annual salary (hours worked * hourly rate). 
The final output should look like: 

"Mary will get a paid trip to Hawaii and a salary of 41600"

"John will get a brand new car and a salary of 93600"... etc.

***Your final output when knitting the file should print all desired outcomes.***


```{r ex4, message = FALSE}



```
### **Exercise 5**

(32 pts) Zodiac signs can be categorized into water signs (cancer, scorpio, pisces), air signs (gemini, libra, aquarius), earth signs (taurus, virgo, capricorn), and fire signs (aries, leo, sagittarius). Create a function such that it's argument is a zodiac sign, and it will return what zodiac category they belong to. Make sure to include in your output the following message "cannot read your input, try again!" in the case that someone types an argument that is not a zodiac sign. 

***Your final output when knitting the file should print the function testing all zodiac signs and something that is is not a zodiac sign***

```{r ex5, message = FALSE}

```