---
title: "Final Exam"
author: "Final Exam - Answer Key"
date: "2026-04-30"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Final Exam DSCI 101 Spring 2026


### **Instructions**

* Below is your final exam for DSCI 101 for Spring 2026.

* **YOU ARE NOT ALLOWED TO SHARE THE CONTENT OF THIS EXAM WITH ANYONE**

* You may consult your class notes or the help buttons in R. 

* You are only allowed to work within RStudio. Any file you need to access, you should do it within RStudio.

* Make sure to answer in full sentences if you see *YOUR WRITTEN ANSWER HERE* after a question. 

* Answer each question using this same file. Your codes should go inside the codechunks. 

* Once you are done with the exam, you must knit the file and load the html document in Sakai. 

### **Dataset Description: Cats!**

The dataset, named `cats_dataset.csv`, contains detailed records of 1000 individual cats. 

Each row represents a unique cat. Below is a description of each column: 

* `breed`: The breed of the cat (there are 30 different breeds). 
* `age`: The age of the cat in years. 
* `weight`: the weight of the cat in kilograms (kg).
* `color`: The color of the cat's coat/fur (there are 15 different colors).
* `gender`: the gender of the cat (Female/Male).
* `purchase_date`: the date the cat was purchased by the owner.


### **File submitted in HTML Format **

Submit the html file in Sakai once you are done with your exam. 


### **Loading Packages**

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(plotly)
```

### **Question 1** (20 points)

a. Read in the dataset `cats_dataset.csv`. Name the dataframe `cats_dataset`. 

```{r question_1a, message = FALSE, warning = FALSE}




```

b. Check the class of the `purchase_date` column in the code chunks. What class should it be? and why is it not reflecting the correct class?  

*YOUR WRITTEN ANSWER HERE*

```{r question_1b, message = FALSE, warning = FALSE}



```

c. Transform the `purchase_date` column into a variable of class "Date" (make this transformation within the same dataframe and make sure the column is still named `purchase_date`). Once you make the transformation, check within the same code chunks the new class of the column `purchase_date`. 

```{r question_1c, message = FALSE, warning = FALSE}



```

d. What is the average weight of cats purchased in the month of "January"? Make sure that the full name of the month (i.e., January) appears on your final result. **Your final output should be a printed dataframe with 1 row and two columns: the month of January and the corresponding average weight of cats purchased in that month.**

```{r question_1d, message = FALSE, warning = FALSE}




```

e. How many cats are from a breed that contains the word "Rex" in it? If there is more than one breed that contains the word "Rex" in it, you need to count how many cats belong to each breed. **Your final output should be a printed dataframe with 2 rows and two columns: the breeds that contain the word "Rex" in it and the number of cats that belong to each breed**

```{r question_1e, message = FALSE, warning = FALSE}



```



### **Question 2** (20 points)

a. Create a new function called `cat_function` such that it's argument is `weight_of_cat`. 

If the weight of the cat is less than or equal to 4, then the function should return "Skinny Cat". 

If the weight of the cat is greater than 4 and less than 7, then the function should return "Average Cat". 

If the weight of the cat is higher than or equal to 7, then the function should return "Cute Chubby Cat". 

**After creating the function, test the function (individually) with weights of 3, 5, and 8.**

```{r question_2a, message = FALSE, warning = FALSE}



```

b. Apply the function you created in part a to the `cats_dataset` dataframe using the `map_chr` function (or any other function in the `map` family). Then calculate the average weight across cats for each of the new weight categories you created. ***Your final output should be a printed dataframe with two columns and three rows that shows you the new categories of weights and average weight of cats in each category.***   

```{r question_2b, message = FALSE, warning = FALSE}



```


### **Question 3** (20 points)

Recreate Plot # 1 - Plotly Graph. Recreating this interactive plot will require data wrangling, pivoting, and proper labeling. Follow the following parts to be able to recreate the plot: 

a. Using the `cats_dataset` dataframe, select columns `breed`, `color`, and `age`. Then filter to keep the rows with colors "Black", "Blue", "Brown", "Orange", and "Red". Save it into a new dataframe. **Do not print the dataframe**

```{r question_3a, message = FALSE, warning = FALSE}



```


b. Using the dataframe created in part a, compute the average age per breed and color. **Do not print the dataframe**  

```{r question_3b, message = FALSE, warning = FALSE}





```

c. Reshape the dataframe from part b into a wide format, such that the first column is `breed`, the second column is `Black`, the third column is `Blue`... etc. The core of the table should reflect the average age for the respective breed and color. **Do not print the dataframe**  

```{r question_3c, message = FALSE, warning = FALSE}



```


d. Using the dataframe created in part c, recreate Plot # 1 (exactly as seen in the picture) using the function `plot_ly.` **Your final output should be a printed interactive plot**

```{r question_3d, message = FALSE, warning = FALSE}




```

### **Question 4** (20 points)

Using the `gender` and `age` columns, take the ages from the males as a sample and perform a bootstrap analysis. Estimate the standard error of the **median**. Perform 1000 iterations.

Note: You can turn a column of a datafame into a vector by using the following code `dataframe$column_name`. **Your final output should be the printed value of the estimated standard error.**

```{r question_4, message = FALSE, warning = FALSE}



```

### **Question 5** (20 points)

In her YouTube account, Dr. Abdalla has a collection of 10 videos of her favorite Dancing with the Stars performances. There are 4 videos by Val, 2 by Derek, 2 by Whitney, 1 by Mark, and 1 by Daniella.      

If you shuffle the videos, the theoretical probability that the video that plays will be a performance by Mark is is `1/10` or `0.1` 

Create a function that simulates selecting a Dancing with the Stars perfomance at random. The arguments of the function should be `i`, `all_videos`, and `pro_name`, representing the iterations, the vector of all videos, and the name of the professional dancer you want to compute the probability on, respectively. 

Use this function along with one of the `map` family functions to simulate the experimental probability of getting a Mark video. Use 1000 iterations. 

**Your final output should be the printed experimental probability of getting a Mark video.**

 
```{r question_5, message = FALSE, warning = FALSE}





```

### **Extra Credit** (5 points)

Open a new R script file and recreate the shiny app EXACTLY as it is. The dataframe used to create the bar graph is `mtcars`. 
Colors are "hotpink", "magenta", and "plum".
If you decide to do this for extra credit, you can upload this R script along with your html file. 

No credit will be give if your app does not run. 

