---
title: "Final Exam"
author: "YOUR NAME HERE"
date: "2025-12-11"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Final Exam DSCI 101 Fall 2025


### **Instructions**

* Below is your final exam for DSCI 101 for Fall 2025.

* **YOU ARE NOT ALLOWED TO SHARE THE CONTENT OF THIS EXAM WITH ANYONE**

* You may consult your class notes or the help buttons in R. 

* You are only allowed to work within RStudio. Any file you need to access, you should do it within RStudio.

* Make sure to answer in full sentences if you see *YOUR WRITTEN ANSWER HERE* after a question. 

* Answer each question using this same file. Your codes should go inside the codechunks. 

* Once you are done with the exam, you must knit the file and load the html document in Sakai. 

### **Dataset Description: Student Performance**

The dataset, named `student_performance.csv`, contains detailed records of the academic performance of 1000 students. 

Each row represents a student. Below is a description of each column: 

* `gender`: the gender of the student (Female/Male). 
* `ethnicity`: the ethnicity of the student. There are 5 categories (African American, Asian, Multiracial, White, and Other. 
* `parental_education`: the highest level of education the student's parents achieved. There are 6 categories (Associate's Degree, Bachelor's Degree, High School, Master's Degree, Some College, and Some High School).
* `math_score`: student's score in a math standardized test.
* `reading_score`: student's score in a reading standardized test.
* `writing_score`: student's score in a writing standardized test.
* `birth_date`: student's birthdate. 

### **File submitted in HTML Format **

Submit the html file in Sakai once you are done with your exam. 


### **Loading Packages**

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(plotly)
```

### **Question 1** (20 points)

a. Read in the dataset `student_performance.csv`. Name the dataframe `student_performance`. 

```{r question_1a, message = FALSE, warning = FALSE}

```

b. Check the class of the `birth_date` column in the code chunks. What class should it be? and why is it not reflecting the correct class?  

*YOUR WRITTEN ANSWER HERE*

```{r question_1b, message = FALSE, warning = FALSE}

```

c. Transform the `birth_date` column into a variable of class "Date" (make this transformation within the same dataframe and make sure the column is still named `birth_date`). Once you make the transformation, check within the same code chunks the new class of the column `birth_date`. 

```{r question_1c, message = FALSE, warning = FALSE}


```

d. Count how many students were born on each day of the week. Make sure that the full name of the week (i.e., Friday, Monday... etc.) appears on your results. **Your final output should be a printed dataframe with 7 rows and two columns: the day of the week and the corresponding number of students.**

```{r question_1d, message = FALSE, warning = FALSE}

```

e. What is the average reading score for each one of the ethnicities? When the final output is printed, each ethnicity must show up in all UPPER CASE. **Your final output should be a printed dataframe with five rows and two columns: the ethnicity in UPPER CASE and the average reading score for each ethnicity.**

```{r question_1e, message = FALSE, warning = FALSE}

```



### **Question 2** (20 points)

a. Create a new function called `education_function` such that it's argument is `education_level`. If the education level is "Associate's Degree", "Bachelor's Degree", "Master's Degree", or "Some College", then the function will return "College and Above". 

If the education level is "Some High School" or "High School", then the function will return "High School or Below". 

After creating the function, test the function (individually) with "Associate's Degree" and "Some High School".

***Note: Make sure that the names of the education level in the function are spelled as mentioned on the instructions above*** 


```{r question_2a, message = FALSE, warning = FALSE}


```

b. Apply the function you created in part a to the `student_performance` dataframe using the `map_chr` function (or any other function in the `map` family). Then calculate how many students per the new education level there are. ***Your final output should be a printed dataframe with two columns and two rows that shows you the new categories of education level and the number of students in each category.***   

```{r question_2b, message = FALSE, warning = FALSE}

```


### **Question 3** (20 points)

Recreate Plot # 1 - Plotly Graph. Recreating this interactive plot will require data wrangling, pivoting, and proper labeling. Follow the following parts to be able to recreate the plot: 

a. Using the `student_performance` dataframe, select columns `ethnicity`, `parental_education`, and `math_score`. Save it into a new dataframe. **Do not print the dataframe**

```{r question_3a, message = FALSE, warning = FALSE}

```


b. Using the dataframe created in part a, compute the average math score per ethnicity and parental education. **Do not print the dataframe**  

```{r question_3b, message = FALSE, warning = FALSE}




```

c. Reshape the dataframe from part b into a wide format, such that the first column is `parental_education`, the second column is `African_American`, the third column is `Asian`... etc. The core of the table should reflect the average math score for the respective degree and ethnicity. **Do not print the dataframe**  

```{r question_3c, message = FALSE, warning = FALSE}

```


d. Using the dataframe created in part c, recreate Plot # 1 (exactly as seen in the picture) using the function `plot_ly.` **Your final output should be a printed interactive plot**

```{r question_3d, message = FALSE, warning = FALSE}



```

### **Question 4** (20 points)

Using the `gender` and `math_score` columns, take the mathematics scores from the females as a sample and perform a bootstrap analysis. Estimate the standard error of the **maximum**. Perform 1000 iterations.

Note: You can turn a column of a datafame into a vector by using the following code `dataframe$column_name`. **Your final output should be the printed value of the estimated standard error.**

```{r question_4, message = FALSE, warning = FALSE}




```

### **Question 5** (20 points)

Suppose Dr. Abdalla's iPod has a total of 10 songs: 4 from Elvis, 3 from Shakira, and 3 from Rihanna.     

If you shuffle the songs, the theoretical probability that the song that will play is an Elvis song is `4/10` or `0.4` 

Create a function that simulates selecting a song at random. The arguments of the function should be `i`, `all_songs`, and `artist_name`, representing the iterations, the vector of all songs, and the name of the artist you want to compute the probability on, respectively. 

Use this function along with one of the `map` family functions to simulate the experimental probability of getting an Elvis song. Use 1000 iterations. 

**Your final output should be the printed experimental probability of getting an Elvis song.**

 
```{r question_5, message = FALSE, warning = FALSE}





```

### **Extra Credit** (5 points)

Open a new R script file and recreate the shiny app EXACTLY as it is. The dataframe used to create the scatter plot is `mtcars`. If you decide to do this for extra credit, you can upload this R script along with your html file. 

No credit will be give if your app does not run. 

